Hidden Talents Multiverse Analysis

### 2021-09-27 manuscript/manuscript.Rmd
update repo for Child Development

---
output:
  bookdown::pdf_document2:
    latex_engine: xelatex
    toc: false
    number_sections: false
    keep_tex: true
    extra_dependencies: ["booktabs","caption","longtable","float"]
geometry: margin=1in
mainfont: Times
fontsize: 11pt
endnote: no
urlcolor: blue
header-includes:
- \usepackage{siunitx}
- \usepackage[utf8]{inputenc}
- \usepackage[english]{babel}
- \setlength{\parindent}{2em}
- \setlength{\parskip}{0em}
- \linespread{1.5}
- \usepackage{titlesec}
- \titleformat{\section}[block]{\normalsize\bfseries\filcenter}{}{1em}{}
- \titleformat{\subsubsection}[block]{\bfseries\itshape}{}{1em}{}
- \usepackage{indentfirst}
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyhf{}
- \fancyhead[R]{\thepage}
- \fancyhead[L]{HIDDEN TALENTS IN CONTEXT}
- \renewcommand\headrulewidth{0pt}
- \usepackage{fixltx2e}
- \usepackage{caption}
- \captionsetup[table]{labelfont={bf},labelformat={default}, labelsep={newline}}
- \captionsetup[figure]{labelfont={bf},labelformat={default}, labelsep={period}}
- \usepackage{lineno}
- \linenumbers
bibliography: references.bib
csl: nature.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = F,
  fig.align = "center",
  fig.pos = "!",  
  fig.show = "asis",
  message = FALSE,
  warning = FALSE
)

@@

\noindent Children developing in harsh conditions often perform poorly on standard cognitive tests. However, the hidden talents approach proposes that performance for some abilities may be enhanced by adversity or improved under ecologically relevant conditions. We examine ecologically relevant content in cognitive testing. Sampling 618 socioeconomically diverse adolescents, we measured performance on two versions of an attention-shifting and working memory updating task---one abstract, the other ecological. Across multiple sources, we measured environmental unpredictability, violence, and poverty, and tested adversity × task version interactions using a preregistered multiverse approach. We did not find adversity × task version interactions for attention shifting, but there were main effects of unpredictability and violence, indicating improved attention shifting irrespective of task version. For working memory updating, violence- and poverty-exposed youth performed worse on the abstract but not the ecological version, equalizing performance among adversity-exposed youth. We found no relation between updating performance and unpredictability.

@@

Children living in harsh conditions tend to perform poorly on many cognitive tests. For example, children raised in poverty often show difficulties learning in school and score lower on standardized tests[@ursache2016a; @farah2006; @noble2007; @hackman2010]. Harsh and unpredictable environments are thought to cause high levels of chronic stress, which impacts brain regions responsible for memory and learning[@blair2012]. Thus, the general consensus is that exposure to adversity undermines cognitive development and leads to a host of cognitive impairments[@hackman2010]. Although it is clear that harsh environments are harmful to children, there is growing interest in the *hidden talents* *approach*, which proposes that some abilities may be enhanced by adversity, especially those abilities that are ecologically relevant to lived experience. Thus, hidden talents are skills and abilities that help children meet real-world challenges in adverse environments[@ellis2017; @ellis2020; @frankenhuis2013; @frankenhuis2020].

The hidden talents framework assumes that development shapes the skills and abilities of individuals in relation to their environment[@ellis2017; @ellis2020; @frankenhuis2013; @frankenhuis2020]. Accordingly, exposure to adversity should shape skills and abilities well-suited for dealing with the challenges associated with harsh environments. However, because specific dimensions of adversity (violence versus resource-scarcity) may pose unique challenges[@ellis2020; @frankenhuis2020; @mclaughlin2019; @sheridan2014], the hidden talents approach proposes that cognitive skills should match specific dimensions of adversity[@frankenhuis2020]. For instance, physically abused children can detect angry facial expressions quickly and accurately[@gibb2009; @pollak2008; @pollak2009], people living in poverty have shown some evidence of enhanced social perception[@piff2018], and anxiously attached people are good at deception detection[@eindor2014; @shoda2013] (see[@ellis2017; @ellis2020] for reviews).

Although some abilities themselves are more ecologically relevant to specific dimensions of adversity, hidden talents may also emerge under ecologically relevant conditions. For example, previous studies of hidden talents have tested cognition under ecologically relevant *contexts and situations*. In particular, studies have experimentally manipulated the current context to be uncertain and mildly stressful (or not) and measured performance on executive functioning and working memory tasks[@mittal2015; @young2018] (but see[@nweze2021] for similar results without uncertainty). Strikingly, people from unpredictable environments showed improved performance on attention-shifting and working memory updating tasks. However, this effect only emerged under conditions of uncertainty. There was no performance improvement under control conditions.

Hidden talents may also emerge when testing *content and stimuli* are ecologically relevant to adversity-exposed individuals. Typical cognitive tests use abstract content that is largely detached from the real world. Such abstract materials may disadvantage youth with limited exposure to formal education. Cultural psychologists and anthropologists have demonstrated this fact across the world. For example, people from low income and non-western cultures tend to perform worse on standard versions of cognitive tests, but this performance gap closes or disappears with more ecologically relevant versions of the same tests[@banerjee2017; @greenfield2014; @ogbu1981; @richardson1991; @schliemann2002]. These findings pose an important question: Can ecologically relevant testing materials improve cognitive performance among adversity-exposed youth?

One way to test this question is to simply replace sterile testing materials with ecologically relevant content. For example, one study tested memory and reasoning among young adults who varied in their exposures to violence[@frankenhuis2020a]. They created two versions of a memory and reasoning task, one neutral (chronological age) and one more ecologically relevant (social dominance). They found that people with more current exposures to violence were equally good, or even better, at memorizing social-dominance relationships compared to people with fewer exposures. Notably, such violence exposures were associated with worse memory for neutral (age) relationships. There were no associations revealed between violence exposures and reasoning performance. Overall, this study suggests that ecologically relevant testing materials may "level the playing field" for adversity-exposed youth[@ellis2020; @frankenhuis2020; @rifkin-graboi2021; @vantassel-baska2018]. In addition, although executive functions and working memory are related to many educational outcomes, it is unclear how schools and teachers might leverage skills that specifically depend on inducing uncertainty. That seems impractical, at best, and unethical, at worst.

@@

We also evaluated the relative robustness of our models to arbitrary data processing decisions using a multiverse analysis approach[@steegen2016; @simonsohn2020]. Specifically, we systematically and explicitly generated all possible combinations of (arbitrary) data processing decisions and analyzed them simultaneously. We conducted a separate multiverse analysis for each task (attention shifting and working memory updating) and each measure of adversity exposure combination. All multiverse analyses used the same underlying statistical models.

@@

@@

## Overview {#overview}

To test our hypotheses, we performed a set of primary and secondary analyses on attention-shifting and working memory performance. Our primary analyses tested the effects of task version, adversity dimensions, and adversity dimension × task version interactions. Specifically, we tested the main effect of environmental unpredictability (confirmatory), violence exposure (confirmatory), and poverty exposure (exploratory), and the interaction between each adversity dimension with task version (abstract vs. ecological) on attention-shifting and working memory performance. We ran separate analyses for each adversity dimension and outcome combination.

Before conducting analyses, we identified six arbitrary data processing decisions each with two alternatives (see [Arbitrary Decisions](#abritrary) section). These criteria were 1) including the after-school club sample or not; 2) including participants with minor distractions during testing or not; 3) including participants who missed one trap question or not (those who missed more than 1 were already excluded); 4) including participants receiving special education (participants receiving more \> 60 minutes/day were already excluded) or not; 5) including participants who performed at or below chance levels of accuracy on the attention-shifting task or not (see [attention shifting data cleaning procedure]( https://osf.io/kvt5e/?view_only=92d55bd3b9834468acd3a7cef5c3febd) for more information); and 6) residualizing social desirability scores out of adversity scores or not. Using all possible combinations of arbitrary decisions, we constructed a multiverse of `r length(multi_data_list)` datasets. The minimum and maximum sample sizes were `r map_dbl(multi_data_list, "n") %>% min()` and `r map_dbl(multi_data_list, "n") %>% max()`, respectively.

We ran the same underlying mixed-effects linear regression model. Analyses were conducted using R[@rcoreteam2020] and mixed models were fit using the lme4 package[@bates2015]. Our mixed-effects models included five fixed terms: the main effect of task version (ecological or abstract version, within-subjects), the main effect of childhood adversity (between-subjects), the interaction between task version and adversity exposure, the main effect of age (control variable), and the main effect of test distraction ratings (control variable). Task version was grand mean centered (abstract versions = -1, ecological versions = 1). All childhood adversity measures were standardized before model fitting, and all models included a random intercept for task version nested in participants.

@@

To facilitate interpretation of multiverse results, we report median standardized $\beta$s, median 95% confidence intervals, and percent significant for each term in our model (see Table \@ref(tab:table2)). Percent significant refers to the number of analyses (out of `r length(multi_data_list)`) that resulted in a significant effect for a particular term and reflects the relative sensitivity (or robustness) of an effect to arbitrary data processing decisions. For the main effects of adversity and task version × adversity interactions, we used a bootstrapping technique[@simonsohn2020; @orben2019; @orben2019a] to compute overall *p*-values, which indicate whether a $\beta$ of a term is larger than expected given the null (the median $\beta$ = 0; see [Statistical Inferences](#inference)).

@@

```{r table1}
table.01.0 %>% 
  kbl(
    align = "lcccccccccc",  
    booktabs = T,
    escape = F,
    caption = "\\emph{Bivariate correlations and descriptive statistics for all measures of adversity and their subcomponents.}"
  ) %>% 
  add_header_above(c(" ","Unpredictability" = 3, "Violence" = 3, "Poverty" = 4)) %>% 
  kable_styling(latex_options = "scale_down") %>% 
  footnote(
    general       = "Column headers denote adversity constructs and rows indicate components. The upper triangle contain sample sizes for each correlation.",
    general_title = "Note:", 
    symbol        = c("p < .05,","p < .01"),
    symbol_manual = c("*", "**"),
    footnote_as_chunk = T,
    title_format = "italic",
    threeparttable = T
  )
```

Figures \@ref(fig:figure2) and \@ref(fig:figure3) visualize multiverse task version × adversity interaction results for attention shifting and working memory performance, including depictions of performance on abstract and ecological task versions across high ($+$ 1 *SD*) and low ($-$ 1 *SD*) adversity exposure (Figures \@ref(fig:figure2)a and \@ref(fig:figure3)a), p-curves associated with each interaction term (Figures \@ref(fig:figure2)b and \@ref(fig:figure3)b), sorted interaction -coefficients across arbitrary decisions (Figures \@ref(fig:figure2)c and \@ref(fig:figure3)c), sample sizes (Figures \@ref(fig:figure2)d and \@ref(fig:figure3)d), and a specification grid showing the arbitrary data processing decisions associated with each effect (Figures \@ref(fig:figure2)e and \@ref(fig:figure3)e).

We preregistered all primary analyses on the Open Science Framework (OSF; see the [osf project link](https://osf.io/6r95z/?view_only=92d55bd3b9834468acd3a7cef5c3febd), including an [updated primary analysis plan](https://osf.io/4vsnz/?view_only=92d55bd3b9834468acd3a7cef5c3febd), which included more details. We conducted secondary analyses after our primary analyses (see [Secondary Analyses](#secondary) below). Before conducting these analyses, we uploaded secondary analysis plans (see the [first secondary analysis plan](https://osf.io/7fu35/?view_only=92d55bd3b9834468acd3a7cef5c3febd) and the [second plan](https://osf.io/wcauf/?view_only=92d55bd3b9834468acd3a7cef5c3febd), explaining the goals and purpose of each secondary analysis. A Red Team Critic[@market] successfully reproduced and verified (see this [red team report](https://osf.io/xv94z/?view_only=92d55bd3b9834468acd3a7cef5c3febd)) all analyses prior to submission.

@@

```{r figure2, fig.height=7.5, fig.width=6.5, fig.cap='(ref:figure2)'}
fig2
```

@@

```{r table2}
table.02.0 %>% 
  ungroup() %>% 
  mutate(iv = factor(iv, c("Unpredictability","Violence","Poverty"))) %>%
  arrange(dv,iv) %>% 
  select(-dv) %>% 
  mutate(across(.cols = contains("p_percent"), ~ str_replace(.x, "%", "\\\\%"))) %>% 
   kable(
    col.names = c(" ","$\\beta$", "95\\% CI","\\emph{p} (\\%)","\\emph{p}", "$\\beta$", "95\\% CI","\\emph{p} (\\%)","\\emph{p}"),
    booktabs = T,
    escape = F,
    caption = "\\emph{Median standardized effects, 95\\% confidence intervals, percent significant, and bootstrapped p-values for the main effect of adversity and interaction terms for adversity by task version for attention shifting and working memory updating.}"
  ) %>% 
  add_header_above(c(" " = 1, "Main Effect" = 4, "Interaction" = 4)) %>% 
  group_rows(group_label = "Attention Shifting", start_row = 1, end_row = 3) %>% 
  group_rows(group_label = "Working Memory Updating", start_row = 4, end_row = 6) %>% 
  kable_styling(latex_options = "scale_down") %>% 
  footnote(
    general = "The p (%) column reflects the number of analyses that produced p-values < .05 out of all analyses for a given multiverse analyses. The total number of analyses for each measure was 64. Overall p-values were computed using a bootstrapping resampling method and reflect the probability of obtaining an effect size as extreme or more extreme given the median effect is 0.",
    general_title = "Note:", 
    footnote_as_chunk = F,
    title_format = "italic",
    threeparttable = T
  )
```

>>>>

---
output:
  bookdown::pdf_document2:
    latex_engine: xelatex
    toc: false
    number_sections: false
    keep_tex: true
    extra_dependencies: ["booktabs","longtable"]
geometry: margin=1in
mainfont: Times
fontsize: 12pt
endnote: no
urlcolor: blue
header-includes:
- \usepackage{float}
- \usepackage{siunitx}
- \usepackage[utf8]{inputenc}
- \usepackage[english]{babel}
- \setlength{\parindent}{2em}
- \setlength{\parskip}{0em}
- \linespread{2}
- \usepackage{titlesec}
- \titleformat{\section}[block]{\normalsize\bfseries\filcenter}{}{}{}
- \titleformat{\subsection}[block]{\normalsize\bfseries}{}{}{}
- \titleformat{\subsubsection}[block]{\bfseries\itshape}{}{}{}
- \titlespacing{\section}{0pt}{0pt}{0pt}
- \titlespacing{\subsection}{0pt}{0pt}{0pt}
- \titlespacing{\subsubsection}{0pt}{0pt}{0pt}
- \usepackage{indentfirst}
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyhf{}
- \fancyhead[R]{\thepage}
- \fancyhead[L]{HIDDEN TALENTS IN CONTEXT}
- \renewcommand\headrulewidth{0pt}
- \usepackage{fixltx2e}
- \usepackage{caption}
- \usepackage{setspace}
- \captionsetup[table]{labelfont={bf},labelformat={default}, labelsep={newline}}
- \captionsetup[figure]{labelfont={bf},labelformat={default}, labelsep={period}}
- \usepackage{lineno}
bibliography: references.bib
csl: apa.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = F,
  fig.align = "center",
  fig.pos = "!t", 
  out.extra = "",
  fig.show = "asis",
  message = FALSE,
  warning = FALSE
)

@@

\noindent Adversity-exposed youth often perform poorly on standard cognitive tests. However, the hidden talents approach proposes some abilities are enhanced by adversity, especially under ecologically relevant conditions. Sampling 618 adolescents (*M~age~* = `r txt1_sam$non_arb$age$mean`, *SD~age~* = `r txt1_sam$non_arb$age$sd`; `r (txt1_sam$non_arb$sex$females/txt1_sam$non_arb$n$total * 100) %>% round(2)`% female; `r (txt1_sam$non_arb$race$non_white/txt1_sam$non_arb$n$total * 100) %>% round(2)`% non-White), we administered two versions of an attention shifting and working memory updating task—one abstract, one ecological. We measured environmental unpredictability, violence, and poverty, and tested adversity × task version interactions. There were no interactions for attention shifting, but there were main effects of unpredictability and violence, indicating improved attention shifting, on average. For working memory updating, violence- and poverty-exposed youth performed worse on the abstract but not the ecological version, equalizing performance. We found no relation between updating and unpredictability.

@@

Children living in harsh conditions tend to perform poorly on many cognitive tests. For example, children raised in poverty often show difficulties learning in school and score lower on standardized tests [@ursache2016a; @farah2006; @noble2007; @hackman2010]. Harsh and unpredictable environments are thought to cause high levels of chronic stress, which impacts brain regions responsible for memory and learning [@blair2012]. Thus, the general consensus is that exposure to adversity undermines cognitive development and leads to a host of cognitive impairments [@hackman2010]. Although it is clear that harsh environments are harmful to children, there is growing interest in the *hidden talents* *approach*, which proposes that some abilities may be enhanced by adversity, especially those abilities that are ecologically relevant to lived experience. Thus, hidden talents are skills and abilities that help children meet real-world challenges in adverse environments [@ellis2017; @ellis2020; @frankenhuis2013; @frankenhuis2020].

The hidden talents framework assumes that development shapes the skills and abilities of individuals in relation to their environment [@ellis2017; @ellis2020; @frankenhuis2013; @frankenhuis2020]. Accordingly, exposure to adversity should shape skills and abilities well-suited for dealing with the challenges associated with harsh environments. However, because specific dimensions of adversity (violence versus resource-scarcity) may pose unique challenges [@ellis2020; @frankenhuis2020; @mclaughlin2019; @sheridan2014], the hidden talents approach proposes that cognitive skills should match specific dimensions of adversity [@frankenhuis2020]. For instance, physically abused children can detect angry facial expressions quickly and accurately [@gibb2009; @pollak2008; @pollak2009], people living in poverty have shown some evidence of enhanced social perception [@piff2018], and anxiously attached people are good at deception detection [@eindor2014; @shoda2013] to list a few examples [see @ellis2017; @ellis2020 for reviews].

Although some abilities themselves are more ecologically relevant to specific dimensions of adversity, hidden talents may also emerge under ecologically relevant conditions. For example, previous studies of hidden talents have tested cognition under ecologically relevant *contexts and situations*. In particular, studies have experimentally manipulated the current context to be uncertain and mildly stressful (or not) and measured performance on executive functioning and working memory tasks [@mittal2015; @young2018; but see @nweze2021 for similar results without uncertainty]. Strikingly, people from unpredictable environments showed improved performance on attention-shifting and working memory updating tasks. However, this effect only emerged under conditions of uncertainty. There was no performance improvement under control conditions.

Hidden talents may also emerge when testing *content and stimuli* are ecologically relevant to adversity-exposed individuals. Typical cognitive tests use abstract content that is largely detached from the real world. Such abstract materials may disadvantage youth with limited exposure to formal education. Cultural psychologists and anthropologists have demonstrated this fact across the world. For example, people from low income and non-western cultures tend to perform worse on standard versions of cognitive tests, but this performance gap closes or disappears with more ecologically relevant versions of the same tests [@banerjee2017; @greenfield2014; @ogbu1981; @richardson1991; @schliemann2002]. These findings pose an important question: Can ecologically relevant testing materials improve cognitive performance among adversity-exposed youth?

One way to test this question is to simply replace sterile testing materials with ecologically relevant content. For example, one study tested memory and reasoning among young adults who varied in their exposures to violence [@frankenhuis2020a]. They created two versions of a memory and reasoning task, one neutral (chronological age) and one more ecologically relevant (social dominance). They found that people with more current exposures to violence were equally good, or even better, at memorizing social-dominance relationships compared to people with fewer exposures. Notably, such violence exposures were associated with worse memory for neutral (age) relationships. There were no associations revealed between violence exposures and reasoning performance. Overall, this study suggests that ecologically relevant testing materials may "level the playing field" for adversity-exposed youth [@ellis2020; @frankenhuis2020; @rifkin-graboi2021; @vantassel-baska2018]. In addition, although executive functions and working memory are related to many educational outcomes, it is unclear how schools and teachers might leverage skills that specifically depend on inducing uncertainty. That seems impractical, at best, and unethical, at worst.

@@

We also evaluated the relative robustness of our models to arbitrary data processing decisions using a multiverse analysis approach [@steegen2016; @simonsohn2020]. Specifically, we systematically and explicitly generated all possible combinations of (arbitrary) data processing decisions and analyzed them simultaneously. We conducted a separate multiverse analysis for each task (attention shifting and working memory updating) and each measure of adversity exposure combination. All multiverse analyses used the same underlying statistical models.

@@

The current study collected data from `r txt1_sam$raw$n$total` adolescents (*M~age~* = `r txt1_sam$raw$age$mean`, *SD~age~* = `r txt1_sam$raw$age$sd`) from two sources across Salt Lake City, Utah. The first was a middle school and the second were 5 after-school clubs. The middle school sample (seventh and eighth graders, *N* = `r txt1_sam$raw$n$wjms`; `r txt1_sam$raw$sex$wjms$females` females) formed `r (txt1_sam$raw$n$wjms/txt1_sam$raw$n$total * 100) %>% round(2)`% of the total sample and is socioeconomically and ethnically diverse (`r (txt1_sam$raw$race$non_white/txt1_sam$raw$n$total * 100) %>% round(2)`% non-White; `r (txt1_sam$raw$race$hispanic/txt1_sam$raw$n$total * 100) %>% round(2)`% Hispanic). According to school district records, `r (txt1_sam$raw$free_lunch$wjms/txt1_sam$raw$n$wjms * 100) %>% round(2)`% of students were receiving economic assistance based on family income and/or homelessness (i.e., free or reduced-price lunch or fee waivers). The clubs (*N* = `r txt1_sam$raw$n$bgc`; `r txt1_sam$raw$sex$bgc$female` females) contained a wider age range (*M~age~*= `r txt1_sam$raw$age$bgc$mean`, *SD~age~* = `r txt1_sam$raw$age$bgc$sd`). `r (txt1_sam$raw$free_lunch$bgc/txt1_sam$raw$n$bgc * 100) %>% round(2)`% were receiving free or reduced-price lunch and were also ethnically diverse.

We measured attention shifting with two versions of the color--shape task [@mittal2015; @miyake2012; @friedman2008], which involves categorizing stimuli according to one of two rules (see below). On each trial, the rule is displayed on the top of the screen. Using arrow keys, participants categorized the stimulus as quickly and accurately as possible. Response options appeared at the bottom of the screen (see Figure \@ref(fig:figure1)).

The ecological attention-shifting task replaced shapes and colors with faces (see Figure \@ref(fig:figure1)b). Faces were either male or female (gender), and were happy or angry (emotion). If a happy female face appeared on screen and the rule was 'gender', for example, participants categorized the face as female. If the rule was 'emotion', participants instead categorized the face as happy. The stimuli for the ecological attention-shifting task were selected from the NimStim library [@tottenham2009]. The same female and male models were used for happy and angry expressions.

To calculate attention-shifting scores, we first cleaned trial-level data [@friedman2008; @miyake2000]. We removed fast and slow reaction times and removed trials after incorrect responses (see [attention shifting data cleaning procedure](https://osf.io/kvt5e/?view_only=92d55bd3b9834468acd3a7cef5c3febd) for details). We then subtracted the average reaction time on valid repeat trials from the average reaction time on valid switch trials to calculate switch cost scores (*M~switch~* - *M~repeat~*). The switch cost measures the relative delay in reaction time caused by switching between categorization rules. Average switch costs on the standard and ecological versions were `r txt2_dvs$abs_shifting_sc$mean` *ms* (*SD* = `r txt2_dvs$abs_shifting_sc$sd` *ms*) and `r txt2_dvs$eco_shifting_sc$mean` ms (*SD* = `r txt2_dvs$eco_shifting_sc$sd` ms), respectively (*r* = `r txt5_cor$shifting$r`, *n* = `r txt5_cor$shifting$n`). Higher switch costs mean longer reaction times on switch trials compared to repeat trials. Larger switch costs indicate less efficient attention shifting.

Working memory updating was measured using two versions of the continuous counters task [@unsworth2015; @unsworth2008; @young2018], in which participants track a sequence of stimuli containing three different objects (see Figure \@ref(fig:figure1)). Participants keep running counts of each object in sequences nine items long. Each object was displayed for one second at a time. At the end of each sequence, participants reported the number of times each object appeared. For each round, there were three possible correct answers, one for each object type.

The third interview variable focused on residential changes from birth to age 13 (reported retrospectively). Specifically, interviewers probed participants about the number of different homes they lived in and, in particular, whether moving to a new home involved living with new or different adults. On average, participants experienced `r txt3_ivs$unp_interview_moves$mean` (*SD* = `r txt3_ivs$unp_interview_moves$sd`) moves that involved changes in residential adults.

The questionnaire included a measure of perceived childhood unpredictability [@mittal2015; @young2018; @szepsenwol2015]. Participants responded to six items that assessed exposure to unpredictability up to age 13: 1) "Things were pretty calm and stable in my house" (reversed); 2) "People moved in and out of my house a lot"; 3) "I had a hard time knowing what my parent(s) or other people in my house were going to say or do from day-to-day"; 4) "My parent(s) argued or fought a lot with each other or with other people"; 5) "My parent(s) changed jobs a lot"; 6) "My parental situation changed a lot (for example, divorce or separation of parents, parents starting new romantic relationships, parents leaving the home)". Participants rated each item on a scale from 1 (never true) to 5 (very often true). Responses to these items were averaged to create a self-reported perceived childhood unpredictability measure (*M* = `r txt3_ivs$unp_perceived$mean`, *SD* = `r txt3_ivs$unp_perceived$sd`, $\alpha$ = `r txt4_rel$unp`). Higher scores indicated more perceived exposure to unpredictable environments.

We measured violence exposure using the Neighborhood Violence Scale [NVS, @frankenhuis2020a; @frankenhuis2018] and two items about witnessing and involvement in physical fights. The NVS contains 7 items rated on a scale from 1 (never true) to 5 (very often true): 1) "I grew up in a safe neighborhood"; 2) "Crime was common in the neighborhood where I grew up"; 3) "In the neighborhood where I grew up, people had plenty of money" (reversed); 4) "In the neighborhood where I grew up, physical fights were common"; 5) "In the neighborhood where I grew up, shootings or stabbings occurred"; 6) "In the neighborhood where I grew up, most people felt unsafe walking alone after dark"; 7) "Where I grew up, it was important to be able to defend yourself against physical harm". The physical fighting items asked about witnessing fights: "Based on your experiences at school and in your neighborhood, how many times did you see or hear someone being beaten up in real life?"; and involvement in fights: "How many times were you in a physical fight at school or in your neighborhood?". Both items were rated on a scale from 1 (0 times) to 8 (12 or more times).

Participants also reported their parents' occupation. Participants described, in general terms, their mother's and father's jobs and the kind of work involved. Trained coders cross-referenced these descriptions with 2010 census job codes [@hout2018]. We calculated mother and father occupational prestige by matching census codes with prestige ratings from an established occupational prestige rating dataset [@hout2016]. Prestige codes ranged from 0-100, with higher scores indicating higher occupational prestige for a particular job code. Mother and father prestige codes were averaged to create a parent occupational prestige score (*M* = `r txt3_ivs$ses_interview_occup$mean`, *SD* = `r txt3_ivs$ses_interview_occup$sd`).

We included two control variables: age and interviewer-rated distractions during testing. We included these because older adolescents tend to score higher on cognitive tests than younger adolescents [@best2010; @best2011; @huizinga2006], and disruptions during the tasks could negatively impact performance. Interviewers completed quantitative ratings of the testing environment including 1) the level of noise; 2) the level of chaos; 3) how often there were interruptions; and 4) how often the participant appeared distracted. Interviewers rated each item on a scale from 1 (not at all) to 7 (much more than average). Items were averaged to create a composite (*M* = `r txt3_ivs$test_envr$mean`, *SD* = `r txt3_ivs$test_envr$sd`), where higher scores indicate more distractions.

We also included the Child Social Desirability (CSD) scale [@miller2014]. The measure includes 14 items such as "Have you ever broken a rule?" and "Do you sometimes feel angry when you don't get your way?". Participants indicated 1 (yes) or 0 (no) for each item. More socially desirable responses were scored as 1. Items were then summed to create a social desirability index (*M* = `r txt3_ivs$csd$mean`, *SD* = `r txt3_ivs$csd$sd` , $\alpha$ = `r txt4_rel$csd`).

@@

## Data Analysis Strategy {#multiverse}

We used a multiverse analysis approach for all analyses to systematically evaluate the robustness (or sensitivity) of analyses across all arbitrary data processing decisions [see  @simonsohn2020; @steegen2016 for details]. We identified six arbitrary data processing decisions. First, we conducted analyses with and without the after-school club sample. We did this because our procedure was slightly different across the two samples and the testing environment was more disruptive in the clubs compared with the middle school. Second, we conducted analyses with and without participants with minor distractions during testing. Third, we analyzed the data with and without participants who missed one trap question (those who missed more than 1 were already excluded). Fourth, we analyzed data with and without participants having any amount of special education (participants receiving more \> 60 minutes were already excluded). Fifth, we analyzed data with and without participants who performed at or below chance levels of accuracy on the attention-shifting task (see [attention shifting data cleaning procedure](https://osf.io/kvt5e/?view_only=92d55bd3b9834468acd3a7cef5c3febd) for more information).

To test our hypotheses, we performed a set of primary and secondary analyses on attention-shifting and working memory performance. Our primary analyses tested the effects of task version, adversity dimensions, and adversity dimension × task version interactions. Specifically, we tested the main effect of environmental unpredictability (confirmatory), violence exposure (confirmatory), and poverty exposure (exploratory), and the interaction between each adversity dimension with task version (abstract vs. ecological) on attention-shifting and working memory performance. We ran separate analyses for each adversity dimension and outcome combination. Using all possible combinations of arbitrary decisions, we constructed a multiverse of `r length(multi_data_list)` datasets. The minimum and maximum sample sizes were `r map_dbl(multi_data_list, "n") %>% min()` and `r map_dbl(multi_data_list, "n") %>% max()`, respectively.

We ran the same underlying mixed-effects linear regression model. Analyses were conducted using R [@rcoreteam2020] and mixed models were fit using the lme4 package [@bates2015]. Our mixed-effects models included five fixed terms: the main effect of task version (ecological or abstract version, within-subjects), the main effect of childhood adversity (between-subjects), the interaction between task version and adversity exposure, the main effect of age (control variable), and the main effect of test distraction ratings (control variable). Task version was grand mean centered (abstract versions = -1, ecological versions = 1). All childhood adversity measures were standardized before model fitting, and all models included a random intercept for task version nested in participants.

@@

To facilitate interpretation of multiverse results, we report median standardized $\beta$s, median 95% confidence intervals, and percent significant for each term in our model (see Table \@ref(tab:table2)). Percent significant refers to the number of analyses (out of `r length(multi_data_list)`) that resulted in a significant effect (two-sided *p*-value \< .05) for a particular term and reflects the relative sensitivity (or robustness) of an effect to arbitrary data processing decisions. For the main effects of adversity and task version × adversity interactions, we used a bootstrapping technique to compute overall *p*-values. The technique makes it possible to know if a multiverse analysis produced a median effect size larger than expected, given the null hypothesis that the median $\beta$ = 0 [see @simonsohn2020 for details].

@@

```{r table1, results='asis'}
table.01.0 %>% 
  kbl(
    align = "lcccccccccc",  
    booktabs = T,
    escape = F,
    caption = "\\emph{Bivariate correlations and descriptive statistics for measures of adversity and their components.}"
  ) %>% 
  add_header_above(c(" ","Unpredictability" = 3, "Violence" = 3, "Poverty" = 4)) %>% 
  kable_styling(latex_options = c("scale_down", "hold_position")) %>% 
  footnote(
    general       = "Column headers denote adversity constructs and rows indicate components. The upper triangle contain sample sizes for each correlation.",
    general_title = "Note:", 
    symbol        = c("p < .05,","p < .01"),
    symbol_manual = c("*", "**"),
    footnote_as_chunk = T,
    title_format = "italic",
    threeparttable = T
  )
```

Figures \@ref(fig:figure2) and \@ref(fig:figure3) visualize multiverse task version × adversity interaction results for attention shifting and working memory performance, including depictions of performance on abstract and ecological task versions across high ($+$ 1 *SD*) and low ($-$ 1 *SD*) adversity exposure (Figures \@ref(fig:figure2)a and \@ref(fig:figure3)a), p-curves associated with each interaction term (Figures \@ref(fig:figure2)b and \@ref(fig:figure3)b), sorted interaction $\beta$-coefficients across arbitrary decisions (Figures \@ref(fig:figure2)c and \@ref(fig:figure3)c), sample sizes (Figures \@ref(fig:figure2)d and \@ref(fig:figure3)d), and a specification grid showing the arbitrary data processing decisions associated with each effect (Figures \@ref(fig:figure2)e and \@ref(fig:figure3)e).

We preregistered all primary analyses on the Open Science Framework (OSF; see the [osf project link](https://osf.io/6r95z/?view_only=92d55bd3b9834468acd3a7cef5c3febd)), including an [updated primary analysis plan](https://osf.io/4vsnz/?view_only=92d55bd3b9834468acd3a7cef5c3febd), which included more details. We conducted secondary analyses after our primary analyses (see [Secondary Analyses](#secondary) below). Before conducting these analyses, we uploaded secondary analysis plans (see the [first secondary analysis plan](https://osf.io/7fu35/?view_only=92d55bd3b9834468acd3a7cef5c3febd) and the [second plan](https://osf.io/wcauf/?view_only=92d55bd3b9834468acd3a7cef5c3febd)), explaining the goals and purpose of each secondary analysis. A Red Team Critic [@redteammarket2021] successfully reproduced and verified (see this [red team report](https://osf.io/xv94z/?view_only=92d55bd3b9834468acd3a7cef5c3febd)) all analyses prior to submission.

@@

```{r figure2, fig.height=7.25, fig.width=6.25, fig.cap='(ref:figure2)'}
fig2
```

@@

```{r table2, results='asis'}
table.02.0 %>% 
  ungroup() %>% 
  mutate(iv = factor(iv, c("Unpredictability","Violence","Poverty"))) %>%
  arrange(dv,iv) %>% 
  select(-dv) %>% 
  mutate(across(.cols = contains("p_percent"), ~ str_replace(.x, "%", "\\\\%"))) %>% 
   kable(
    col.names = c(" ","$\\beta$", "95\\% CI","\\emph{p} (\\%)","\\emph{p}", "$\\beta$", "95\\% CI","\\emph{p} (\\%)","\\emph{p}"),
    booktabs = T,
    escape = F,
    caption = "\\emph{Median standardized effects, 95\\% confidence intervals, percent significant, and bootstrapped p-values for the main effect of adversity and interaction terms for adversity by task version for attention shifting and working memory updating.}"
  ) %>% 
  add_header_above(c(" " = 1, "Main Effect" = 4, "Interaction" = 4)) %>% 
  group_rows(group_label = "Attention Shifting", start_row = 1, end_row = 3) %>% 
  group_rows(group_label = "Working Memory Updating", start_row = 4, end_row = 6) %>% 
  kable_styling(latex_options = "scale_down") %>% 
  footnote(
    general = "The p (%) column reflects the number of analyses that produced p-values < .05 out of all analyses for a given multiverse analyses. The total number of analyses for each measure was 64. Overall p-values were computed using a bootstrapping resampling method and reflect the probability of obtaining an effect size as extreme or more extreme given the median effect is 0.",
    general_title = "Note:", 
    footnote_as_chunk = T,
    title_format = "italic",
    threeparttable = T
  )
```
